---
title: "Assessment_1B_bco7000"
author: "BrookeKneebush"
date: '2022-03-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RMarkdown editing

**R Markdown** is a file format for making dynamic documents with R. An **R Markdown** document is written in markdown (*an easy-to-write plain text format*) and contains chunks of embedded R code, like the document below.

This [LINK](https://rmarkdown.rstudio.com/lesson-8.html) provides information on RMarkdown basics.

```{r}
library(tidyverse)
library(janitor)
library(skimr)

fastfood <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')

```


```{r}

fastfood%>%head(10)

```

```{r}

knitr::kable(fastfood[1:10,1:5])

```
Arranging observations
top_n()
```{r}

fastfood%>% top_n(10)

```

8. Display the observations that has more than 1000 calories

```{r}

fastfood%>%filter(calories>1000)

```

9. Arrange observations with more than 40 in total_fat and more than 80 in total_carb in the descending order and save them to a new variable (dataset) called `dont_eat_this`

```{r}

dont_eat_this<-fastfood%>%filter(total_fat>40 & total_carb>80)%>%arrange(desc(total_fat))

```

10. Using the initial dataset variable, use `case_when()` to create a new variable `heavy_food` which is equal to "heavy" when total calories are greater than 500, "low" when total calories are less than 250 and "average" for all other cases. Count the number of observations in each created category.

```{r}

heavy_food<-fastfood%>%mutate(heavy_food=case_when(calories>500~"heavy",calories<250~"low",T~"average"))

heavy_food%>%
  count(heavy_food, sort=TRUE)


```

11. Display the types of variables in the dataset using `skimr` package 

```{r}

library(skimr)

fastfood%>%skim()

```

12. Present the count observations from each restaurant in a descending order

```{r}

fastfood%>%count(restaurant)%>%arrange(desc(n))

```

Show the number of distnct items on a menu in the dataset

```{r}

item<-fastfood %>%
  distinct(item)

```

13. Using groupings (group_by()), summarise and display the average number of calories for each restaurant.

```{r}

fastfood%>%group_by(restaurant)%>%summarise(restaurant_average_calories=mean(calories, na.rm=TRUE))%>%ungroup()

```

14. Add variables to the dataset, which:
-  calculates the average calories per type of restaurant and call it `average_calories` `

```{r}

average_calories<-fastfood%>%mutate(average_calories=mean(calories, na.rm=TRUE))%>%group_by(restaurant)%>%summarise(average_calories=mean(calories, na.rm=TRUE))%>%ungroup()

```

14. Add variables to the dataset, which:
-  calculates the maximum total_fat per type of restaurant and call it `max_fat` 

```{r}

fastfood%>%mutate(max_fat=max(total_fat, na.rm=TRUE))%>%group_by(restaurant)%>%summarise(max_fat=max(total_fat, na.rm=TRUE))

```

14. Add variables to the dataset, which:
calculates the minimum cholesterol per type of restaurant and call it `min_cholesterol` 

```{r}

fastfood%>%mutate(min_cholesterol=min(cholesterol, na.rm=TRUE))%>%group_by(restaurant)%>%summarise(min_cholesterol=min(min_cholesterol, na.rm=TRUE))

```

15. Display the data vis of total fat per each type of restaurant. Write a narration (2-3 sentences) why you believe this type of data viz presents such information

I believe this type of data viz (bar chart geom_bar) is an effective method of presenting the total fat per each type of restaurant, because the height of the bars clearly represents the values of total_fat for each type of restaurant.  Each of the restaurants is represented on the x axis and total_fat per each restaurant is represented on the y axis.   Of the four options available for one discrete and one continuous variable, in this instance, the geom_col presents the data in the most easy to read fashion.

```{r}


data(fastfood)
fastfood%>% ggplot(aes(restaurant, total_fat)) + geom_col()

```

16. Add a variable to the dataset, which calculates the sum of cholesterol and sodium and call it `cholesterol_sodium`.

```{r}

fastfood%>%mutate(cholesterol_sodium=cholesterol+sodium)

```

Remove the variable `salad`

```{r}

fastfood%>%select(-salad)

```


17. Use observations for Mcdonalds to plot sugar variable against protein with `geom_point()` 

```{r}

fastfood%>% ggplot(aes(sugar, protein )) + geom_point()

```

